---
title: "simpleRegress"
author: "Riley Wheeler"
date: "Summer 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26

```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

### CHEST

```{r}
chesty <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chesty)
summary.lm(chesty)
```

The adjusted r-squared fo this model (CHEST) is .79; the model for NECK had an adjusted R-squared of .75. 
So far, the best model is the one based on the CHEST measurements.

```{r}
new <- data.frame(CHEST=c(17,22,26))
predict(chesty,new)
predict(chesty,new,se.fit=TRUE)
```


### WEIGHT


```{r}
weighty <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weighty)
summary.lm(weighty)
```

The adjusted r-squared fo this model (WEIGHT) is .74, which is less than the .75 of neck.
So far, the best model is the one based on the CHEST measurements.

```{r}
new <- data.frame(WEIGHT=c(17,22,26))
predict(weighty,new)
predict(weighty,new,se.fit=TRUE)
```


### HEADWTH


```{r}
headwdthy <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwdthy)
summary.lm(headwdthy)
```

The adjusted r-squared fo this model (HEADWTH) is .73, less than .75 from the neck.
So far, the best model is the one based on the CHEST measurements.

```{r}
new <- data.frame(HEADWTH=c(17,22,26))
predict(headwdthy,new)
predict(headwdthy,new,se.fit=TRUE)
```


### HEADLEN


```{r}
healdeny <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(healdeny)
summary.lm(healdeny)
```

The adjusted r-squared fo this model (HEADLEN) is .85, which takes the throne so far as the highest.
So far, the best model is the one based on the Head Length (HEADLEN) measurements.

```{r}
new <- data.frame(HEADLEN=c(17,22,26))
predict(healdeny,new)
predict(healdeny,new,se.fit=TRUE)
```




### SEX


```{r}
sexy <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
abline(sexy)
summary.lm(sexy)
```

The adjusted r-squared fo this model (SEX) is -.01, meaning it has no correlation with length whatsoever.
So far, the best model is the one based on the HEADLEN measurements.

```{r}
new <- data.frame(SEX=c(17,22,26))
predict(sexy,new)
predict(sexy,new,se.fit=TRUE)
```




### MONTH


```{r}
monthy <- lm(LENGTH~MONTH,data=bears)
plot(LENGTH~MONTH,data=bears)
abline(monthy)
summary.lm(monthy)
```

The adjusted r-squared fo this model (MONTH) is .01, meaning this, just like the sex, has no correlation with length
So far, the best model is the one based on the HEADLEN measurements.

```{r}
new <- data.frame(MONTH=c(17,22,26))
predict(monthy,new)
predict(monthy,new,se.fit=TRUE)
```




### AGE


```{r}
agey <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(agey)
summary.lm(agey)
```

The adjusted r-squared fo this model (AGE) is .5; the model for NECK had an adjusted R-squared of .75. 
So far, the best model is the one based on the HEADLEN measurements.

```{r}
new <- data.frame(AGE=c(17,22,26))
predict(agey,new)
predict(agey,new,se.fit=TRUE)
```

# RESULTS

While the NECK had an adjusted R-squared value of .75, the HEADLEN measurement came out with a value of .85. The HEADLEN value has the most significant impact on overall length, while the MONTH and SEX have the lowest correlations.