---
title: "simpleRegress"
author: "Riley Wheeler"
date: "Summer 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26

```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

# Results

## Graphical and Numerical Descriptive Results  

### CHEST  
 
```{r}
chesty <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chesty)
summary.lm(chesty)
```

The adjusted r-squared fo this model (CHEST) is .79; the model for NECK had an adjusted R-squared of .75. 
So far, the best model is the one based on the CHEST measurements. The graph here seems to begin at 40cm length (y intercept), then makes a slow linear ascent through the data points.

```{r}
new <- data.frame(CHEST=c(17,22,26))
predict(chesty,new)
predict(chesty,new,se.fit=TRUE)
```


### WEIGHT


```{r}
weighty <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weighty)
summary.lm(weighty)
```

The adjusted r-squared fo this model (WEIGHT) is .74, which is less than the .75 of neck.
So far, the best model is the one based on the CHEST measurements. The graph here seems to begin at 45cm length (y intercept), then makes a slow linear ascent through the data points.

```{r}
new <- data.frame(WEIGHT=c(17,22,26))
predict(weighty,new)
predict(weighty,new,se.fit=TRUE)
```


### HEADWTH


```{r}
headwdthy <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwdthy)
summary.lm(headwdthy)
```

The adjusted r-squared fo this model (HEADWTH) is .73, less than .75 from the neck.
So far, the best model is the one based on the CHEST measurements. The graph here seems to begin at 40cm length (y intercept), then makes a slow linear ascent through the data points up to 10cm headwth.

```{r}
new <- data.frame(HEADWTH=c(17,22,26))
predict(headwdthy,new)
predict(headwdthy,new,se.fit=TRUE)
```


### HEADLEN


```{r}
healdeny <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(healdeny)
summary.lm(healdeny)
```

The adjusted r-squared fo this model (HEADLEN) is .85, which takes the throne so far as the highest.
So far, the best model is the one based on the Head Length (HEADLEN) measurements. The graph here seems to begin at 40cm length (y intercept), then makes a slow linear ascent through the data points up to 16cm headlen.

```{r}
new <- data.frame(HEADLEN=c(17,22,26))
predict(healdeny,new)
predict(healdeny,new,se.fit=TRUE)
```




### SEX


```{r}
sexy <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
abline(sexy)
summary.lm(sexy)
```

The adjusted r-squared fo this model (SEX) is -.01, meaning it has no correlation with length whatsoever.
So far, the best model is the one based on the HEADLEN measurements. The graph here seems to begin at 60cm length (y intercept), then makes a slight linear decent , almost zero slope, through the data points down to 2.0.

```{r}
new <- data.frame(SEX=c(17,22,26))
predict(sexy,new)
predict(sexy,new,se.fit=TRUE)
```




### MONTH


```{r}
monthy <- lm(LENGTH~MONTH,data=bears)
plot(LENGTH~MONTH,data=bears)
abline(monthy)
summary.lm(monthy)
```

The adjusted r-squared fo this model (MONTH) is .01, meaning this, just like the sex, has no correlation with length
So far, the best model is the one based on the HEADLEN measurements. The graph here seems to begin at 55cm length (y intercept), then makes a very slight linear ascent, almost zero slope, through the data points from January to December.

```{r}
new <- data.frame(MONTH=c(17,22,26))
predict(monthy,new)
predict(monthy,new,se.fit=TRUE)
```




### AGE


```{r}
agey <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(agey)
summary.lm(agey)
```

The adjusted r-squared fo this model (AGE) is .5; the model for NECK had an adjusted R-squared of .75. 
So far, the best model is the one based on the HEADLEN measurements. The graph here seems to begin at 50cm length (y intercept), then makes a somewhat rapid linear ascent through the data points up to the near 200 year old outlier.

```{r}
new <- data.frame(AGE=c(17,22,26))
predict(agey,new)
predict(agey,new,se.fit=TRUE)
```

# Discussion and Conclusion

While the NECK had an adjusted R-squared value of .75, the HEADLEN measurement came out with a value of .85. The HEADLEN value has the most significant impact on overall length, while the MONTH and SEX have the lowest correlations. The Graphs all have a line of best fir inserted with a y-intercept from 40-60. Most of the graphs have a slight positive linear slope; however, MONTH and SEX graphs both have a near-zero slope.


